---
title: "DS5110:Group-Project"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	message = FALSE,
	warning = FALSE,
	echo = TRUE
)
```

## Importing packages

```{r}
library(dplyr)
library(readr)
library(ggplot2)
library(tidyr)
library(gridExtra)
library(grid)
library(stringr)
library(tidyverse)
```

## Importing Data

```{r}

d <- "Datasets"
advCourse <- read_csv(file.path(d,"AdvancedCourseCompletion.csv")) 
ap_part <- read_csv(file.path(d,"ap_participation.csv")) 
ap_perf <- read_csv(file.path(d,"ap_performance.csv"))
art <- read_csv(file.path(d,"artcourse.csv")) 
attendance <- read_csv(file.path(d,"attendance.csv"))
attrition <- read_csv(file.path(d,"AttritionReport.csv")) 

classSizeByClass <- read_csv(file.path(d,"ClassSizebyGenPopulation.csv")) 
classSizeByRace <- read_csv(file.path(d,"ClassSizebyRaceEthnicity.csv"))

dropOut <- read_csv(file.path(d,"dropout.csv")) 

eduAge <- read_csv(file.path(d,"EducatorsbyAgeGroupsReport.csv"))  
enrollByGrade <- read_csv(file.path(d,"enrollmentbygrade.csv"))

gradeStaff <- read_csv(file.path(d,"gradestaffing.csv")) 
gradRate <- read_csv(file.path(d,"gradrates.csv")) 
college <- read_csv(file.path(d,"Gradsattendingcollege.csv")) 
mobilityRate <- read_csv(file.path(d,"mobilityrates.csv"))

StudReten <- read_csv(file.path(d,"retention2021.csv")) 

sat <- read_csv(file.path(d,"sat_performance.csv")) 
selectPop <- read_csv(file.path(d,"selectedpopulations.csv")) 

daysMissed <- read_csv(file.path(d,"ssdr_days_missed.csv"))
eduGen <- read_csv(file.path(d,"staffracegender.csv")) 
staffReten<- read_csv(file.path(d,"staffingretention.csv"))
discipline <- read_csv(file.path(d,"StudentDisciplineDataReport.csv"))

teachData <- read_csv(file.path(d,"teacherdata.csv"))
teachProg <- read_csv(file.path(d,"Teacherprogramarea.csv"))
teacherSalary <- read_csv(file.path(d,"TeacherSalaries.csv"))


```
## EDA

### 1. Number of students who took one or more Advanced Placement exams.
```{r}
ap_part2 <- ap_part |> 
  rename(`1`=`One Test` ,`2`=`Two Tests`,`3`=`Three Tests`
                             , `4`=`Four Tests`,`>=5`=`Five or More Tests`) |>
  select(!c(`Tests Taken`,`Tests Takers`)) |>
  filter(`District Code`!="00000000")

pivot_longer(ap_part2, cols=c(`1`,`2`,`3`,`4`,`>=5`), names_to = "TestsTaken", values_to ="TestTakers") |>  
  group_by(TestsTaken) |> 
  summarise(TestTakers=sum(TestTakers, na.rm=TRUE)) |>
  ggplot( mapping=aes(x=`TestsTaken`,y=`TestTakers`)) +
  geom_bar(stat = "identity") +
  labs(title="Students taking Advanced Placement Exams",
       x="Number of Tests Taken", y="Number of Students")
```

It shows that most of the students took placements exams **only once** while only less that 2500 students had to take 5 or more exams.

### 2. Percentage of tests taken by students with each possible score on the Advanced Placement exam.
```{r}
ap_perf <- ap_perf |>
  select(!c(`% Score 1-2`,`% Score 3-5`)) |>
  filter(`District Code`!=0)
ap_perf <- ap_perf |> rename(`1`=`Score=1` ,`2`=`Score=2`,`3`=`Score=3`
                             , `4`=`Score=4`,`5`=`Score=5`)

pivot_longer(ap_perf, cols=c(`1`,`2`,`3`,`4`,`5`),
             names_to = "Scores",
             values_to ="# of Tests") |>
  group_by(`Scores`) |>
  summarise(`Total`=sum(`# of Tests`, na.rm=TRUE)) |>
  ggplot( mapping=aes(x=`Scores`,y=`Total`)) +
  geom_bar(stat = "identity") +
  labs(title="Scores in Advanced Placement Exams",
       x="Scores Earned", y="Number of tests")
```
Above plot shows that most of the tests had a score of 3 while least number of tests had a score of 5. Since most of the students had taken the exam only once, it could be possible that most of them had a score of 3.

### 3. The rate of Grade 11 and 12 students completing advanced courses by subject area.
```{r}
advCourse <- advCourse %>% 
  select(!c(`Ch 74 Secondary Cooperative Program`,`# Students Completing Advanced`)) |>
  filter(`District Code`!="00000000") 
advCourse |> ggplot( mapping=aes(`% Students Completing Advanced`)) +
  geom_boxplot() +
  labs(title="Rate of students completing advanced courses")
```

It shows that for some of the districts, all students completed the advanced courses while minimum rate was around 25% students. On an average, more than 62.5% of the students in 305 district entries were able to complete the courses.

Districts in which 100% or 0% students completed advanced courses . 
```{r}
advCourse |> select(`District Name`,`District Code`,
                    `% Students Completing Advanced`,
                    `# Grade 11 and 12 Students`) |> 
  filter(`% Students Completing Advanced` == 100.0 |
           `% Students Completing Advanced` == 0 ) |>
  rename(`% Completion` =`% Students Completing Advanced`,`Student Count`=`# Grade 11 and 12 Students`)

```
It can be noticed that the districts with 0% completion has quite low number of students in grade 11 and 12 and except **Martha's Vineyard Charter (District)** and **Pioneer Valley Chinese Immersion Charter (District)** all other schools with 100% completion has comparatively high student count.

Plotting subject-wise Completions.
```{r}
advCourse100 <- advCourse %>%
  rename(`% Science and Technology`=`% Science and Technology...8`,
         `% Computer Science` = `% Science and Technology...9`) %>%
  pivot_longer( cols=c(`% ELA`,`% Math`,`% Science and Technology`,
                               `% Computer Science`,
                             `% History and Social Sciences`,`% Arts`,
                             `% All Other Subjects`),
             names_to = "Subject",
             values_to ="% Completion") |>
  mutate(Subject=gsub("%","",Subject)) |>
  filter(`% Students Completing Advanced` == 100.0)

advCourse100 |>
  ggplot( mapping=aes(x=factor(`Subject`),y=`% Completion`)) +
  geom_boxplot() +
  labs(title="Subject-wise completion among 100% completion rate districts",
       x="Subjects", y="Percentage of Completion") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.1, hjust=1))
```

Clearly, the highest completion has been for Math while surprisingly for Computer Science, it has been the least.

```{r}
advCourse100|>
  ggplot( mapping=aes(x=factor(`District Code`),y=`% Completion`,
                      fill = `Subject`))+
  geom_bar(position="dodge", stat="identity")+
  theme_minimal()+
  theme(legend.position = "bottom") +
  labs(title="District-wise Subject completion among 100% completion districts",
       x="Districts", y="% Completion",
       fill="Subjects")
```

For district code **4660000**, only math has been completed and which is why it seems to join the 100% completion club!
Since, we noticed before, it has only 22 students which is really less compared to rest of the other schools that had 100% completion rate. District code **4970000**, has done really well as it only had 68 students as compared to district code **4890000**, that had 407 students!

### 4. Dropouts
```{r}
dropOut |> rename(`# Enrolled 9-12` =`# Enrolled Grades 09 through 12`,
                  `Dropout`=`# Dropout All Grades`) |>
 select(`District Code`,`# Enrolled 9-12`,`Dropout`,`District Name`) |> drop_na() |>
  filter(`District Code`!="00000000") |>
  filter(`Dropout`== max(Dropout))
```
Boston has the maximum number of dropouts!

### Analyzing data with the responses.

## Tran's work starts here.

```{r}
eduGen <- eduGen %>%
  rename(
    'District Code' = 'District/School Code')

# daysMissed <- daysMissed %>%
#   rename(
#     'District_code' = 'District Code')
# 
# selectPop <- selectPop %>%
#   rename(
#     'District_code' = 'District Code')
# 
# 
# mobilityRate <- mobilityRate %>%
#   rename(
#     'District_code' = 'District Code')
```

# 1. Teacher Salary vs. Graduation Rate
```{r}


gradRate_teacherSalary <- left_join(gradRate, teacherSalary, by="District Code")

gradRate_teacherSalary <- gradRate_teacherSalary %>%
  rename(
    'percent_graduated' = '% Graduated')

gradRate_teacherSalary <- gradRate_teacherSalary %>%
  rename(
    'average_salary' = 'Average Salary')

# print(gradRate_teacherSalary)
```


```{r}
gradRate_teacherSalary_graph <-
  ggplot(gradRate_teacherSalary, aes(x = average_salary, 
                                 y = percent_graduated )) + geom_point() + 
  theme_minimal() + 
  ggtitle("Graduate rate percentage vs. 
          Average teacher salary") + geom_smooth(method=lm) + ylim(50,100)

gradRate_teacherSalary_graph
```


# 2. Teacher Data vs. Graduation Rate
```{r}


gradRate_teacherData <- left_join(gradRate, teachData, by="District Code")

gradRate_teacherData <- gradRate_teacherData %>%
  rename(
    'percent_graduated' = '% Graduated')

gradRate_teacherData <- gradRate_teacherData %>%
  rename(
    'experienced_teacher_percent' = 'Percent of Experienced Teachers')

gradRate_teacherData <- gradRate_teacherData %>%
  rename(
    'student_teacher_ratio' = 'Student / Teacher Ratio')

# print(gradRate_teacherData)
```

```{r}


gradRate_teacherData <- gradRate_teacherData %>%
      mutate_at("student_teacher_ratio", str_replace, "to 1", "")

# print(gradRate_teacherData)
```


```{r}
gradRate_teacherData_graph <-
  ggplot(gradRate_teacherData, aes(x = experienced_teacher_percent, 
                                 y = percent_graduated )) + geom_point() + 
  theme_minimal() + 
  ggtitle("Graduate rate percentage vs. 
          Experienced teacher percentage") + geom_smooth(method=lm) + 
  ylim(75, 100)

gradRate_teacherData_graph
```

Observation: there is a correlation; the more experienced teacher, the higher graduation rate. 

```{r}
StudentTeacherRatio_graph <-
  ggplot(gradRate_teacherData, aes(x = student_teacher_ratio, 
                                 y = percent_graduated )) + geom_point() + 
  theme_minimal() + 
  ggtitle("Graduate rate percentage vs. 
          Student teacher ratio") + geom_smooth(method=lm) 

StudentTeacherRatio_graph
```

# 3. Student Discipline vs. Graduation Rate
```{r}


gradRate_discipline <- left_join(gradRate, discipline, by="District Code")
gradRate_discipline <- na.omit(gradRate_discipline)

gradRate_discipline <- gradRate_discipline %>%
  rename(
    'percent_graduated' = '% Graduated')

gradRate_discipline <- gradRate_discipline %>%
  rename(
    'percent_suspension' = '% In-School Suspension')

# print(gradRate_discipline)
```

```{r}
gradRate_discipline_graph <-
  ggplot(gradRate_discipline, aes(x = percent_suspension, 
                                 y = percent_graduated )) + geom_point() + 
  theme_minimal() + 
  ggtitle("Graduate rate percentage vs. 
          Student discipline") + geom_smooth(method=lm) 

gradRate_discipline_graph
```

# 4. Demographic vs. Graduation Rate
```{r}


gradRate_demographic <- left_join(gradRate, eduGen, by="District Code")
gradRate_demographic <- na.omit(gradRate_demographic)

gradRate_demographic <- gradRate_demographic %>%
  rename(
    'percent_graduated' = '% Graduated')

gradRate_demographic <- gradRate_demographic %>%
  rename(
    'African_American' = 'African American (#)')
gradRate_demographic <- gradRate_demographic %>%
  rename(
    'White' = 'White (#)')

gradRate_demographic <- gradRate_demographic %>%
  rename(
    'Asian' = 'Asian (#)')

gradRate_demographic <- gradRate_demographic %>%
  rename(
    'Hispanic' = 'Hispanic (#)')

gradRate_demographic <- gradRate_demographic %>%
  rename(
    'Native_American' = 'Native American (#)')

gradRate_demographic <- gradRate_demographic %>%
  rename(
    'Hawaiian_Pacific_Islander' = 'Native Hawaiian, Pacific Islander (#)')

gradRate_demographic <- gradRate_demographic %>%
  rename(
    'Multi_race_non_Hispanic' = 'Multi-Race,Non-Hispanic (#)')

gradRate_demographic_long <- pivot_longer(gradRate_demographic, cols=11:17, 
                                          names_to = "Race", 
                                          values_to = "Race_number")

print(gradRate_demographic_long)
```
```{r}
gradRate_demographic_graph <-
  ggplot(gradRate_demographic_long, aes(x = Race_number, 
                                 y = percent_graduated, color = Race)) + 
  geom_point() + 
  theme_minimal() + 
  ggtitle("Graduate rate percentage vs. 
          Staff Demographic") + xlim(0, 1000)+ facet_grid(Race ~.) + 
  geom_smooth(method=lm) 

gradRate_demographic_graph 
```

# 5. Staffing Retention vs. Graduation Rate

```{r}

gradRate_staffingRetention <- left_join(gradRate, staffReten, 
                                        by="District Code")
gradRate_staffingRetention <- na.omit(gradRate_staffingRetention)

print(gradRate_staffingRetention)
```

# 6. Graduation Rate vs. Day missed

```{r}

daysMissed$District_code <- str_pad(daysMissed$`District Code`, 8, pad = "0")

# gradRateRename <- gradRate %>%
#   rename(
#     'District_code' = 'District Code')

print(daysMissed)

```
```{r}
gradRate_daysMissed <- left_join(gradRate, daysMissed, by="District Code")
gradRate_daysMissed <- na.omit(gradRate_daysMissed)

gradRate_daysMissed <- gradRate_daysMissed %>%
  rename(
    'percent_graduated' = '% Graduated')

gradRate_daysMissed_long <- pivot_longer(gradRate_daysMissed, cols=13:17, 
                                         names_to = "Num_days_missed", 
                                         values_to = "percent_num_days_missed")

print(gradRate_daysMissed_long)

```

```{r}
gradRate_daysMissed_graph <-
  ggplot(gradRate_daysMissed_long, aes(x = percent_num_days_missed, 
                                 y = percent_graduated, 
                                 color = Num_days_missed)) + geom_point() + 
  theme_minimal() + 
  ggtitle("Graduate rate percentage vs. 
          SAT") + facet_grid(Num_days_missed ~.) + geom_smooth(method=lm)
gradRate_daysMissed_graph 
```

# 7. SAT vs. Graduation Rate

```{r}

gradRate_SAT <- left_join(gradRate, sat, by="District Code")


gradRate_SAT <- gradRate_SAT %>%
  rename(
    'percent_graduated' = '% Graduated')

gradRate_SAT <- gradRate_SAT %>%
  rename(
    'tests_taken' = 'Tests Taken')

gradRate_SAT_long <- pivot_longer(gradRate_SAT, cols=12:14, 
                                  names_to = "SAT_test_types", 
                                  values_to = "SAT_test_scores")

gradRate_SAT_long <- gradRate_SAT_long[!(is.na(gradRate_SAT_long$SAT_test_scores)), ]

print(gradRate_SAT_long)
```

```{r}
gradRate_SAT_graph <-
  ggplot(gradRate_SAT_long, aes(x = SAT_test_scores, 
                                 y = percent_graduated, 
                                color = SAT_test_types)) + geom_point() + 
  theme_minimal() + 
  ggtitle("Graduate rate percentage vs. 
          SAT") + facet_grid(SAT_test_types ~.) + geom_smooth(method=lm) + ylim(60,100)
gradRate_SAT_graph 
```

# 8. Graduation plan Vs. Graduation rate

```{r}

# gradRate_gradplan <- left_join(gradRate, plansforHSgrad, by="District Code")
# 
# gradRate_gradplan <- gradRate_gradplan %>%
#   rename(
#     'percent_graduated' = '% Graduated')
# 
# gradRate_gradplan_long <- pivot_longer(gradRate_gradplan, cols=11:20, names_to = "Plan_type", values_to = "Plan_percentage")
# 
# print(gradRate_gradplan_long)
# use cmd/ctrl + shift + c to uncomment
```
```{r}
# gradRate_gradPlan_graph <-
#   ggplot(gradRate_gradplan_long, aes(x = Plan_percentage, 
#                                  y = percent_graduated, color = Plan_type)) + geom_point() + 
#   theme_minimal() + 
#   ggtitle("Graduate rate percentage vs. 
#           Plan after high school") + facet_grid(Plan_type ~.) + geom_smooth(method=lm) 
# gradRate_gradPlan_graph 
```


# 9. Graduation rate vs. Students Background

```{r}
print(selectPop)

# gradRateRename1 <- gradRate %>%
#   rename(
#     'District_code' = 'District Code')

selectPop$District_code <- str_pad(selectPop$`District Code`, 8, pad = "0")

print(selectPop)

```
```{r}
gradRate_selectPop <- left_join(gradRate, selectPop, by="District Code")

gradRate_selectPop <- gradRate_selectPop %>%
  rename(
    'percent_graduated' = '% Graduated')

gradRate_selectPop <- gradRate_selectPop %>%
  rename(
    'economically_disadvantaged' = 'Economically Disadvantaged %')


print(gradRate_selectPop)

```
```{r}
gradRate_selectPop_graph <-
  ggplot(gradRate_selectPop, aes(x = economically_disadvantaged, 
                                 y = percent_graduated)) + geom_point() + 
  theme_minimal() + 
  ggtitle("Graduate rate percentage vs. 
          Economically Disadvantaged % Students") 
gradRate_selectPop_graph 
```

```{r}

print(mobilityRate)

# gradRateRename2 <- gradRate %>%
#   rename(
#     'District_code' = 'District Code')

mobilityRate$District_code <- str_pad(mobilityRate$`District Code`, 8, pad = "0")

print(mobilityRate)


```
```{r}
gradRate_mobilityRate <- left_join(gradRate, mobilityRate, by="District Code")

gradRate_mobilityRate <- gradRate_mobilityRate %>%
  rename(
    'percent_graduated' = '% Graduated')

gradRate_mobilityRate_long <- pivot_longer(gradRate_mobilityRate, cols=12:13, 
                                           names_to = "churn_intake", 
                                          values_to = "churn_intake_percentage")

print(gradRate_mobilityRate_long)

```
```{r}
gradRate_mobility_graph <-
  ggplot(gradRate_mobilityRate_long, aes(x = churn_intake_percentage, 
                                 y = percent_graduated, 
                                 color = churn_intake)) + geom_point() + 
  theme_minimal() + 
  ggtitle("Graduate rate percentage vs. 
          Mobility rate") + facet_grid(churn_intake ~.) + xlim(0, 15) + 
  ylim(50, 100) +  geom_smooth(method=lm) 
gradRate_mobility_graph 
```

## Tran's work end here

## Cleaning Data
```{r}
sat <- sat %>% mutate(`Total Score` = `Reading / Writing` + Math) %>% 
  select(!Writing)

enrollByGrade <- enrollByGrade %>% 
  mutate(`HS Enrollment` = `9` + `10` + `11` + `12`) %>% 
  select(`District Code`, `HS Enrollment`, Total) %>% 
  rename(Enrollment = Total)

ap_part <- ap_part %>% select(`District Code`, `Tests Takers`)

staffReten <- staffReten %>% select(`District Code`, `Teacher % Retained`) %>% 
  rename(`Teacher Retention Rate` = `Teacher % Retained`)

classSize <- inner_join(classSizeByClass,classSizeByRace) %>%
  select(!c(`Number of Students`,`Total # of Classes`,
                        `District Name`,`English Language Learner %`,
            `Students with Disabilities %`,`Economically Disadvantaged %`
            ))

college <- college %>% 
  rename(`Percent Going to College`=`Attending Coll./Univ. (%)`) %>%
  select(`District Code`, `Percent Going to College`)
  
attendance <- attendance %>% select(`District Code`, `Attendance Rate`, 
                                    `Average # of Absences`)
attrition <- attrition %>% select(`District Code`, ALL) %>% 
  rename(Attrition = ALL)

advCourse <- advCourse %>% 
  select(`District Code`, `% Students Completing Advanced`, `% Math`, 
         `% ELA`) %>% rename(`Adv Course % Math` = `% Math`, 
                             `Adv Course % ELA` = `% ELA`)

dropOut <- dropOut |> select(`District Code`,`% Dropout All Grades`)

gradRate <- gradRate %>% select(`District Code`, `% Graduated`, `% Dropped Out`)

art <- art %>% 
  mutate(`% in an Art Course` = `All Grades` / `Total Students` * 100) %>% 
  select(`District Code`, `% in an Art Course`)

eduAge <- eduAge %>% 
  mutate(`% of Teachers <40` = (`<26 yrs (# )` + `26-32 yrs (#)` + 
                                  `33-40 yrs (#)`) / `FTE Count` * 100) %>% 
  select(`District Code`, `% of Teachers <40`)

discipline <- discipline %>% 
  mutate(`% Disciplined` = `Students Disciplined` / `Students` * 100) %>% 
  select(`District Code`, `% Disciplined`)

convertPerc <- function(x, na.rm = TRUE) format(round((x / eduGen$`FTE Count`) 
                                                      * 100, 3), nsmall = 3)
eduGen <- eduGen %>% mutate_at(c("Females (#)", "African American (#)",
                                 "Asian (#)","Hispanic (#)","White (#)",
                                 "Native American (#)",
                                 "Native Hawaiian, Pacific Islander (#)",
                                 "Multi-Race,Non-Hispanic (#)", "Males (#)"),
                               convertPerc) %>%
  rename(`% Female Teachers`="Females (#)",`% African American Teachers`=
           "African American (#)",`% Asian Teachers`="Asian (#)",
         `% Hispanic Teachers`="Hispanic (#)",`% White Teachers`="White (#)",
         `% Native American Teachers`="Native American (#)",
         `% Native Hawaiian, Pacific Islander Teachers`=
           "Native Hawaiian, Pacific Islander (#)",
         `% Multi-Race,Non-Hispanic Teachers`="Multi-Race,Non-Hispanic (#)",
         `% Male Teachers`="Males (#)") %>%
  select(!c(`District/School Name`,`FTE Count`))

mobilityRate <- mobilityRate %>% select(!c(`District Name`,
                                           `Churn/Intake Enroll`,
                                           `Stability Enroll`))
  
teachData <- teachData %>% select(!`District Name`)
teachData$`Student / Teacher Ratio` <- substr(
  teachData$`Student / Teacher Ratio`,1,
  nchar(teachData$`Student / Teacher Ratio`)-5) %>% parse_number()

selectPop <- selectPop %>% rename(`High Needs %`=`High Needs #...16`) %>%
  select(`District Code`,`First Language Not English %`,
         `English Language Learner %`,`Students With Disabilities %`,
         `High Needs %`,`Economically Disadvantaged %`)

```

## Joining all tables
```{r}
eduData <- inner_join(sat, enrollByGrade, by = "District Code") %>% 
  inner_join(ap_part, by = "District Code") %>% 
  inner_join(staffReten, by = "District Code") %>% 
  inner_join(classSize, by = "District Code") %>% 
  inner_join(college, by = "District Code") %>% 
  inner_join(attendance, by = "District Code") %>%
  inner_join(attrition, by = "District Code") %>%
  inner_join(advCourse, by = "District Code") %>% 
  inner_join(gradRate, by = "District Code") %>% 
  inner_join(art, by = "District Code") %>% 
  inner_join(eduAge, by = "District Code") %>% 
  inner_join(discipline, by = "District Code") %>%
  inner_join(eduGen, by = "District Code") %>%
  inner_join(teachData, by = "District Code") %>%
  inner_join(dropOut, by = "District Code") %>%
  inner_join(mobilityRate, by = "District Code") %>%
  inner_join(selectPop, by = "District Code") %>%
  mutate(`Percent of HS in AP` = `Tests Takers` / `HS Enrollment` * 100) %>% 
  mutate(`Adjusted Score` = `Total Score` * `% Graduated` / 100) %>% drop_na()

```

```{r}
summary(eduData)
```
**Inference from summary: **
1) Neither Reading/Writing, nor Math has a perfect score in SAT, same goes for the total score.

2) Among races, at least one district had Hispanic and White students in domination,even though the third quartile of Hispanic students is at 16.43%.

3) Even though on average Male % students is more than Female % among districts, the minimum and maximum of gender % is higher for Females.

4) At least one district/school has 100% graduate rate and 0% drop rate and yet maximum percentage of students going to College is only 91.70.

5) At least one district/school has % Students Completing Advanced as 100% and none has 100% attendance rate.

5) The school/district with minimum % of Teachers Licensed, has almost half of the teachers unlicensed.

6) None of the school/district has **only** experienced teachers, and in at least one school, 67% teachers are not experienced.

7) The data is taken for the COVID time-period(2020-21), yet at least one school had Percent Teaching In-Field as 100%.

8) Even though the schools are in a country where English is the most-commonly spoken language, at least one school has 83.400% students whose first Language is not English.

9) None of the schools has 0% of High Needs or Economically Disadvantaged students.


To see if genders had a relation with Total SAT score.
```{r}
g1 <- eduData %>% ggplot( mapping=aes(x=`Female %`,y=`Total Score`))+
  geom_point() +
  geom_smooth(method=lm) +
  labs(title="Positive Relationship:Female% vs SAT",
       x="Female % ", y="SAT Score") +
  theme_minimal()

g2 <- eduData %>% ggplot( mapping=aes(x=`Male %`,y=`Total Score`))+
  geom_point() +
  geom_smooth(method=lm) +
  labs(title="Negative Relationship:Male% vs SAT ",
       x="Male % ", y="SAT Score") +
  theme_minimal()

gridExtra::grid.arrange(
  g1, g2,nrow=1 ,top = textGrob("Relationship of SAT score with gender",
                                   gp=gpar(fontsize=15,font=3)))
```

## Average Class Size vs SAT, Graduate Rate, drop rate and Enrollment in college

```{r}
g1 <- eduData %>% ggplot( mapping=aes(x=`Average Class Size`,y=`Total Score`))+
  geom_point() +
  geom_smooth(method=lm) +
  theme_minimal()

g2 <- eduData %>% ggplot( mapping=aes(x=`Average Class Size`,
                                      y=`Percent Going to College`))+
  geom_point() +
  geom_smooth(method=lm) +
  theme_minimal()

g3 <- eduData %>% ggplot( mapping=aes(x=`Average Class Size`,
                                      y=`% Dropped Out`))+
  geom_point() +
  geom_smooth(method=lm) +
  theme_minimal()

g4 <- eduData %>% ggplot( mapping=aes(x=`Average Class Size`,y=`% Graduated`))+
  geom_point() +
  geom_smooth(method=lm) +
  theme_minimal()

gridExtra::grid.arrange(
  g1, g2,g3,g4 ,
  top = textGrob("Relationship of responses with Average Class Size",
                                   gp=gpar(fontsize=15,font=3)))

```
There happens to be an outlier, which is in fact a correctly reported value (~45), so we won't remove it. Average Class Size has negative relationship with SAT score and % graduated, while it has a positive relationship with % going to college and %dropped out. 


```{r}

g1 <- eduData %>% ggplot( mapping=aes(x=`% Students Completing Advanced`,
                                      y=`Total Score`))+
  geom_point() +
  geom_smooth(method=lm) +
  theme_minimal()

g2 <- eduData %>% ggplot( mapping=aes(x=`% Students Completing Advanced`,
                                      y=`Percent Going to College`))+
  geom_point() +
  geom_smooth(method=lm) +
  theme_minimal()

g3 <- eduData %>% ggplot( mapping=aes(x=`% Students Completing Advanced`,
                                      y=`% Dropped Out`))+
  geom_point() +
  geom_smooth(method=lm) +
  theme_minimal()

g4 <- eduData %>% ggplot( mapping=aes(x=`% Students Completing Advanced`,
                                      y=`% Graduated`))+
  geom_point() +
  geom_smooth(method=lm) +
  theme_minimal()

gridExtra::grid.arrange(
  g1, g2,g3,g4 ,
  top = textGrob("Relationship of responses with % Students Completing Advanced",
                                   gp=gpar(fontsize=15,font=3)))


```
All responses had positive relationship with % Students Completing Advanced except %dropped out, which is expected since it's inverse of %graduated so, here onwards we will consider only one of them.

## Attendance Rate vs Responses
```{r}
g1 <- eduData %>% ggplot( mapping=aes(x=`Attendance Rate`,
                                      y=`Total Score`))+
  geom_point() +
  geom_smooth(method=lm) +
  theme_minimal()

g2 <- eduData %>% ggplot( mapping=aes(x=`Attendance Rate`,
                                      y=`Percent Going to College`))+
  geom_point() +
  geom_smooth(method=lm) +
  theme_minimal()

g3 <- eduData %>% ggplot( mapping=aes(x=`Attendance Rate`,
                                      y=`% Graduated`))+
  geom_point() +
  geom_smooth(method=lm) +
  theme_minimal()

gridExtra::grid.arrange(
  g1,g2,g3,
  top = textGrob("Relationship of responses with Attendance Rate",
                                   gp=gpar(fontsize=15,font=3)))


```
```{r}
g1 <- eduData %>% ggplot( mapping=aes(x=`Average # of Absences`,
                                      y=`Total Score`))+
  geom_point() +
  geom_smooth(method=lm) +
  theme_minimal()

g2 <- eduData %>% ggplot( mapping=aes(x=`Average # of Absences`,
                                      y=`Percent Going to College`))+
  geom_point() +
  geom_smooth(method=lm) +
  theme_minimal()

g3 <- eduData %>% ggplot( mapping=aes(x=`Average # of Absences`,
                                      y=`% Graduated`))+
  geom_point() +
  geom_smooth(method=lm) +
  theme_minimal()

gridExtra::grid.arrange(
  g1,g2,g3,
  top = textGrob("Negative Relationship of responses with Average 
                 Number of Absences", gp=gpar(fontsize=15,font=3)))


```


```{r}
g1 <- eduData %>% ggplot( mapping=aes(x=`Attrition`,
                                      y=`Total Score`))+
  geom_point() +
  geom_smooth(method=lm) +
  theme_minimal()

g2 <- eduData %>% ggplot( mapping=aes(x=`Attrition`,
                                      y=`Percent Going to College`))+
  geom_point() +
  geom_smooth(method=lm) +
  theme_minimal()

g3 <- eduData %>% ggplot( mapping=aes(x=`Attrition`,
                                      y=`% Graduated`))+
  geom_point() +
  geom_smooth(method=lm) +
  theme_minimal()

gridExtra::grid.arrange(
  g1,g2,g3,
  top = textGrob("Negative Relationship of responses with Attrition", 
                 gp=gpar(fontsize=14,font=3)))
```
Even though, all of them has negative relationship, the slopes are different i.e., graduation rate drops with larger difference as compared to other responses.


## Student Background vs Graduation Rate 
```{r}
g1 <- eduData %>% ggplot( mapping=aes(x=`African American %`,
                                      y=`% Graduated`))+
  geom_point() +
  geom_smooth(method=lm) +
  theme_minimal()#-

g2 <- eduData %>% ggplot( mapping=aes(x=`Asian %`,
                                      y=`% Graduated`))+
  geom_point() +
  geom_smooth(method=lm) +
  theme_minimal()

g3 <- eduData %>% ggplot( mapping=aes(x=`Hispanic %`,
                                      y=`% Graduated`))+
  geom_point() +
  geom_smooth(method=lm) +
  theme_minimal()#-

g4 <- eduData %>% ggplot( mapping=aes(x=`White %`,
                                      y=`% Graduated`))+
  geom_point() +
  geom_smooth(method=lm) +
  theme_minimal()

g5 <- eduData %>% ggplot( mapping=aes(x=`Native American %`,
                                      y=`% Graduated`))+
  geom_point() +
  geom_smooth(method=lm) +
  theme_minimal()

g6 <- eduData %>% ggplot( mapping=aes(x=`Native Hawaiian, Pacific Islander %`,
                                      y=`% Graduated`))+
  geom_point() +
  geom_smooth(method=lm) +
  theme_minimal()#-

g7 <- eduData %>% ggplot( mapping=aes(x=`Multi-Race, Non-Hispanic %`,
                                      y=`% Graduated`))+
  geom_point() +
  geom_smooth(method=lm) +
  theme_minimal()

gridExtra::grid.arrange(
  g1,g2,g3,g4,g5,g6,g7,
  top = textGrob("Relationship of responses with % Graduated", 
                 gp=gpar(fontsize=15,font=3)))
```
While "African American", "Hispanic" and "Native Hawaiian,Pacific Islander" students have negative relationship with graduation rate, "Multi-Race, Non-Hispanic" students have somewhat constant graduation rate. Something else to notice is that population of "white" with higher graduation rate is closer to 100% while for others,higher graduation rate is closer to 0%.

```{r}
g1 <- eduData %>% ggplot( mapping=aes(x=`African American %`,
                                      y=`Percent Going to College`))+
  geom_point(alpha=0.1) +
  geom_smooth(method=lm) +
  labs(y="Going College(%)") +
  theme_minimal()

g2 <- eduData %>% ggplot( mapping=aes(x=`Asian %`,
                                      y=`Percent Going to College`))+
  geom_point(alpha=0.1) +
  geom_smooth(method=lm) +
  labs(y="Going College(%)") +
  theme_minimal()

g3 <- eduData %>% ggplot( mapping=aes(x=`Hispanic %`,
                                      y=`Percent Going to College`))+
  geom_point(alpha=0.1) +
  geom_smooth(method=lm) +
  labs(y="Going College(%)") +
  theme_minimal()#-

g4 <- eduData %>% ggplot( mapping=aes(x=`White %`,
                                      y=`Percent Going to College`))+
  geom_point(alpha=0.1) +
  geom_smooth(method=lm) +
  labs(y="Going College(%)") +
  theme_minimal()

g5 <- eduData %>% ggplot( mapping=aes(x=`Native American %`,
                                      y=`Percent Going to College`))+
  geom_point(alpha=0.1) +
  geom_smooth(method=lm) +
  labs(y="Going College(%)") +
  theme_minimal()#-

g6 <- eduData %>% ggplot( mapping=aes(x=`Native Hawaiian, Pacific Islander %`,
                                      y=`Percent Going to College`))+
  geom_point(alpha=0.1) +
  geom_smooth(method=lm) +
  labs(y="Going College(%)") +
  theme_minimal()

g7 <- eduData %>% ggplot( mapping=aes(x=`Multi-Race, Non-Hispanic %`,
                                      y=`Percent Going to College`))+
  geom_point(alpha=0.1) +
  geom_smooth(method=lm) +
  labs(y="Going College(%)") +
  theme_minimal()

gridExtra::grid.arrange(
  g1,g2,g3,g4,g5,g6,g7,
  top = textGrob("Relationship of responses with Percent Going to College", 
                 gp=gpar(fontsize=15,font=3)))
```
When plotted against "Going to College(%)",the only differences were that "African American" showed no significant change,and "Native American" had a negative relationship,while "Multi-Race, Non-Hispanic" and "Native Hawaiian, Pacific Islander" students have slightly positive relationship.

## % of Teachers <40 and responses
```{r}

g1 <- eduData %>% ggplot( mapping=aes(x=`% of Teachers <40`,
                                      y=`Total Score`))+
  geom_point() +
  geom_smooth(method=lm) +
  theme_minimal()

g2 <- eduData %>% ggplot( mapping=aes(x=`% of Teachers <40`,
                                      y=`Percent Going to College`))+
  geom_point() +
  geom_smooth(method=lm) +
  theme_minimal()

g3 <- eduData %>% ggplot( mapping=aes(x=`% of Teachers <40`,
                                      y=`% Graduated`))+
  geom_point() +
  geom_smooth(method=lm) +
  theme_minimal()

gridExtra::grid.arrange(
  g1,g2,g3,
  top = textGrob("Negative Relationship of responses with Teacher's Age", 
                 gp=gpar(fontsize=15,font=3)))

```
Schools with more % of teachers's age less than 40, had a negative affect on Total SAT score and graduation rate while a positive affect on percentage of students going to college. 

```{r}

g1 <- eduData %>% ggplot( mapping=aes(x=`% in an Art Course`,
                                      y=`Total Score`))+
  geom_point() +
  geom_smooth(method=lm) +
  theme_minimal()

g2 <- eduData %>% ggplot( mapping=aes(x=`% in an Art Course`,
                                      y=`Percent Going to College`))+
  geom_point() +
  geom_smooth(method=lm) +
  theme_minimal()

g3 <- eduData %>% ggplot( mapping=aes(x=`% in an Art Course`,
                                      y=`% Graduated`))+
  geom_point() +
  geom_smooth(method=lm) +
  theme_minimal()

gridExtra::grid.arrange(
  g1,g2,g3,
  top = textGrob("Relationship of responses with % in an Art Course", 
                 gp=gpar(fontsize=15,font=3)))

```

```{r}

g1 <- eduData %>% ggplot( mapping=aes(x=`First Language Not English %`,
                                      y=`% Graduated`))+
  geom_point() +
  geom_smooth(method=lm) +
  theme_minimal()

g2 <- eduData %>% ggplot( mapping=aes(x=`English Language Learner %`,
                                      y=`% Graduated`))+
  geom_point() +
  geom_smooth(method=lm) +
  theme_minimal()

g3 <- eduData %>% ggplot( mapping=aes(x=`Students With Disabilities %`,
                                      y=`% Graduated`))+
  geom_point() +
  geom_smooth(method=lm) +
  theme_minimal()

g4 <- eduData %>% ggplot( mapping=aes(x=`High Needs %`,
                                      y=`% Graduated`))+
  geom_point() +
  geom_smooth(method=lm) +
  theme_minimal()

gridExtra::grid.arrange(
  g1,g2,g3,g4,
  top = textGrob("Negative Relationship of % Graduated with different classes%", 
                 gp=gpar(fontsize=15,font=3)))
# English Language Learner %
# Students With Disabilities %
# High Needs %
# First Language Not English %
```
### Percent Teaching In-Field vs 3 responses
```{r}
# Percent Teaching In-Field
g1 <- eduData %>% ggplot( mapping=aes(x=`Percent Teaching In-Field`,
                                      y=`Total Score`))+
  geom_point(alpha=0.1) +
  geom_smooth(method=lm) +
  theme_minimal()

g2 <- eduData %>% ggplot( mapping=aes(x=`Percent Teaching In-Field`,
                                      y=`Percent Going to College`))+
  geom_point(alpha=0.1) +
  geom_smooth(method=lm) +
  theme_minimal()

g3 <- eduData %>% ggplot( mapping=aes(x=`Percent Teaching In-Field`,
                                      y=`% Graduated`))+
  geom_point(alpha=0.1) +
  geom_smooth(method=lm) +
  theme_minimal()

gridExtra::grid.arrange(
  g1,g2,g3,
  top = textGrob("Relationship of responses with % Teaching In-Field", 
                 gp=gpar(fontsize=14,font=3)))
```
Percentage going to college has significantly small negative relationship.

