---
title: "DS5110:Project"
author: "Pratyasha Sharma"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	message = FALSE,
	warning = FALSE,
	echo = TRUE
)
```

## Importing packages

```{r}
library(dplyr)
library(readr)
library(ggplot2)
library(tidyr)
```

## Importing Data

```{r}

d <- "finalProject"
advCourse <- read_csv(file.path(d,"AdvancedCourseCompletion.csv"))
ap_part <- read_csv(file.path(d,"ap_participation.csv"))
ap_perf <- read_csv(file.path(d,"ap_performance.csv"))
art <- read_csv(file.path(d,"artcourse.csv"))
attendance <- read_csv(file.path(d,"attendance.csv"))
attrition <- read_csv(file.path(d,"AttritionReport.csv"))

classSizeByClass <- read_csv(file.path(d,"ClassSizebyGenPopulation.csv"))
classSizeByRace <- read_csv(file.path(d,"ClassSizebyRaceEthnicity.csv"))

dropOut <- read_csv(file.path(d,"dropout.csv"))

eduAge <- read_csv(file.path(d,"EducatorsbyAgeGroupsReport.csv"))
enrollByGrade <- read_csv(file.path(d,"enrollmentbygrade.csv"))
enrollByRaceNgender <- read_csv(file.path(d,"enrollmentbyracegender.csv"))

gradeStaff <- read_csv(file.path(d,"gradestaffing.csv"))
gradRate <- read_csv(file.path(d,"gradrates.csv"))

mobilityRate <- read_csv(file.path(d,"mobilityrates.csv"))

StudReten <- read_csv(file.path(d,"retention2021.csv"))

sat <- read_csv(file.path(d,"sat_performance.csv"))
selectPop <- read_csv(file.path(d,"selectedpopulations.csv"))
daysMissed <- read_csv(file.path(d,"ssdr_days_missed.csv"))
eduGen <- read_csv(file.path(d,"staffracegender.csv"))
discipline <- read_csv(file.path(d,"StudentDisciplineDataReport.csv"))

teachData <- read_csv(file.path(d,"teacherdata.csv"))
teachProg <- read_csv(file.path(d,"Teacherprogramarea.csv"))


```
## EDA

### 1. Number of students who took one or more Advanced Placement exams.
```{r}
ap_part <- ap_part |> 
  rename(`1`=`One Test` ,`2`=`Two Tests`,`3`=`Three Tests`
                             , `4`=`Four Tests`,`>=5`=`Five or More Tests`) |>
  select(!c(`Tests Taken`,`Tests Takers`)) |>
  filter(`District Code`!=0)

pivot_longer(ap_part, cols=c(`1`,`2`,`3`,`4`,`>=5`), names_to = "TestsTaken", values_to ="TestTakers") |>  
  group_by(TestsTaken) |> 
  summarise(TestTakers=sum(TestTakers, na.rm=TRUE)) |>
  ggplot( mapping=aes(x=`TestsTaken`,y=`TestTakers`)) +
  geom_bar(stat = "identity") +
  labs(title="Students taking Advanced Placement Exams",
       x="Number of Tests Taken", y="Number of Students")
```

It shows that most of the students took placements exams only once while only less that 2500 students had to take 5 or more exams.

### 2. Percentage of tests taken by students with each possible score on the Advanced Placement exam.
```{r}
ap_perf <- ap_perf |> 
  select(!c(`% Score 1-2`,`% Score 3-5`)) |>
  # select(`District Code`,`District Name`,`% Score 1-2`,`% Score 3-5`) |>
  filter(`District Code`!=0) |>
ap_perf <- ap_perf |> rename(`1`=`Score=1` ,`2`=`Score=2`,`3`=`Score=3`
                             , `4`=`Score=4`,`5`=`Score=5`)

pivot_longer(ap_perf, cols=c(`1`,`2`,`3`,`4`,`5`), 
             names_to = "Scores", 
             values_to ="# of Tests") |>  
  group_by(`Scores`) |>
  summarise(`Total`=sum(`# of Tests`, na.rm=TRUE)) |>
  ggplot( mapping=aes(x=`Scores`,y=`Total`)) +
  geom_bar(stat = "identity") +
  labs(title="Scores in Advanced Placement Exams",
       x="Scores Earned", y="Number of tests")
```
Above plot shows that most of the tests had a score of 3 while least number of tests had a score of 5. Since most of the students had taken the exam only once, it could be possible that most of them had a score of 3.

### 3. The rate of Grade 11 and 12 students completing advanced courses by subject area.
```{r}
advCourse <- advCourse %>% 
  select(!c(`Ch 74 Secondary Cooperative Program`,`# Students Completing Advanced`)) |>
  filter(`District Code`!=0) 
advCourse |> ggplot( mapping=aes(`% Students Completing Advanced`)) +
  geom_boxplot() +
  labs(title="Rate of students completing advanced courses")
```

It shows that for some of the districts all students completed the advanced courses while minimum rate was around 25% students. On an average, more than 62.5% of the students in 305 district entries were able to complete the courses.

Districts in which 100% or 0% students completed advanced courses . 
```{r}
advCourse |> select(`District Name`,`District Code`,
                    `% Students Completing Advanced`,
                    `# Grade 11 and 12 Students`) |> 
  filter(`% Students Completing Advanced` == 100.0 |
           `% Students Completing Advanced` == 0 ) |>
  rename(`% Completion` =`% Students Completing Advanced`,`Student Count`=`# Grade 11 and 12 Students`)

```
It can be noticed that the districts with 0% completion has quite low number of students in grade 11 and 12 and except **Martha's Vineyard Charter (District)** and **Pioneer Valley Chinese Immersion Charter (District)** all other schools with 100% completion has comparatively high student count.

Plotting subject-wise Completions.
```{r}
advCourse100 <- pivot_longer(advCourse, cols=c(`% ELA`,`% Math`,`% Science and Technology`,
                               `% Computer Science`,
                             `% History and Social Sciences`,`% Arts`,
                             `% All Other Subjects`), 
             names_to = "Subject", 
             values_to ="% Completion") |> 
  mutate(Subject=gsub("%","",Subject)) |>
  filter(`% Students Completing Advanced` == 100.0)

advCourse100 |> 
  ggplot( mapping=aes(x=factor(`Subject`),y=`% Completion`)) +
  geom_boxplot() +
  labs(title="Subject-wise completion among 100% completion rate districts",
       x="Subjects", y="Percentage of Completion") +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.1, hjust=1))
```

Clearly, the highest completion has been for Math while surprisingly for Computer Science, it has been the least.

```{r}
advCourse100|> 
  ggplot( mapping=aes(x=factor(`District Code`),y=`% Completion`,
                      fill = `Subject`))+
  geom_bar(position="dodge", stat="identity")+
  theme_minimal()+
  theme(legend.position = "bottom") +
  labs(title="District-wise Subject completion among 100% completion districts",
       x="Districts", y="% Completion", 
       fill="Subjects") 
```

For district code **4660000**, only math has been completed and which is why it seems to join the 100% completion club!
Since, we noticed before, it has only 22 students which is really less compared to rest of the other schools that had 100% completion rate. District code **4970000**, has done really well as it only had 68 students as compared to district code **4890000**, that had 407 students!

### 4. Dropouts
```{r}
dropOut |> rename(`# Enrolled 9-12` =`# Enrolled Grades 09 through 12`,
                  `Dropout`=`# Dropout All Grades`) |>
 select(`District Code`,`# Enrolled 9-12`,`Dropout`,`District Name`) |> drop_na() |>
  filter(`District Code`!=0) |>
  filter(`Dropout`== max(Dropout))
```




### 4. Class size based on different classes. 
```{r}
classSizeLong <- pivot_longer(classSizeByClass, cols=c(`Female %`,`Male %`,`English Language Learner %`,
                               `Students with Disabilities %`,
                             `Economically Disadvantaged %`),names_to = "Class",
             values_to = "% of Class Size" ) |>
  mutate(Class=gsub("%","",Class))

classSizeLong |> group_by(`Class`) |> filter(Class %in% c(`Female`,`Male`)) |>
  summarise(`Total`=sum(`% of Class Size`, na.rm=TRUE))
  ggplot( mapping=aes(x=factor(`Subject`),y=`% Completion`)) +
  geom_bar() +
  
  labs(title="Subject-wise completion among 100% completion rate districts",
       x="Subjects", y="Percentage of Completion") +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.1, hjust=1))
 
```





