---
title: "group-project"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r}
library(stringr)
library(readr)
library(tidyverse)
library(ggplot2)
library(tidyr)

```

```{r}

mobilityRate <- read_csv(file.path("mobilityrates.csv"))
studentExpenditures <- read_csv(file.path("PerPupilExpenditures.csv"))
plansforHSgrad <- read_csv(file.path("plansforhsgrad.csv"))
  
StudReten <- read_csv(file.path("retention2021.csv"))
sat <- read_csv(file.path("sat_performance.csv"))
selectPop <- read_csv(file.path("selectedpopulations.csv"))
daysMissed <- read_csv(file.path("ssdr_days_missed.csv"))

staffingRetention <-read_csv(file.path("staffingretention.csv"))
eduGen <- read_csv(file.path("staffracegender.csv"))
discipline <- read_csv(file.path("StudentDisciplineDataReport.csv"))

teachData <- read_csv(file.path("teacherdata.csv"))
teachProg <- read_csv(file.path("Teacherprogramarea.csv"))
teacherSalary <-read_csv(file.path("TeacherSalaries.csv"))




eduGen <- eduGen %>%
  rename(
    'District Code' = 'District/School Code')

daysMissed <- daysMissed %>%
  rename(
    'District_code' = 'District Code')

print(gradRate)


print(mobilityRate)
print(studentExpenditures)
print(plansforHSgrad)
print(StudReten)
print(sat)
print(selectPop)
print(daysMissed)



print(staffingRetention)
print(eduGen)
print(discipline)
print(teachData)
print(teachProg)
print(teacherSalary)
```

# Teacher Salary vs. Graduation Rate
```{r}


gradRate_teacherSalary <- left_join(gradRate, teacherSalary, by="District Code")

gradRate_teacherSalary <- gradRate_teacherSalary %>%
  rename(
    'percent_graduated' = '% Graduated')

gradRate_teacherSalary <- gradRate_teacherSalary %>%
  rename(
    'average_salary' = 'Average Salary')

print(gradRate_teacherSalary)
```


```{r}
gradRate_teacherSalary_graph <-
  ggplot(gradRate_teacherSalary, aes(x = average_salary, 
                                 y = percent_graduated )) + geom_point() + 
  theme_minimal() + 
  ggtitle("Graduate rate percentage vs. 
          Average teacher salary") + geom_smooth(method=lm) + ylim(50,100)

gradRate_teacherSalary_graph
```


# Teacher Data vs. Graduation Rate
```{r}


gradRate_teacherData <- left_join(gradRate, teachData, by="District Code")

gradRate_teacherData <- gradRate_teacherData %>%
  rename(
    'percent_graduated' = '% Graduated')

gradRate_teacherData <- gradRate_teacherData %>%
  rename(
    'experienced_teacher_percent' = 'Percent of Experienced Teachers')

gradRate_teacherData <- gradRate_teacherData %>%
  rename(
    'student_teacher_ratio' = 'Student / Teacher Ratio')

print(gradRate_teacherData)
```

```{r}


gradRate_teacherData <- gradRate_teacherData %>%
      mutate_at("student_teacher_ratio", str_replace, "to 1", "")

print(gradRate_teacherData)
```


```{r}
gradRate_teacherData_graph <-
  ggplot(gradRate_teacherData, aes(x = experienced_teacher_percent, 
                                 y = percent_graduated )) + geom_point() + 
  theme_minimal() + 
  ggtitle("Graduate rate percentage vs. 
          Experienced teacher percentage") + geom_smooth(method=lm) + 
  ylim(75, 100)

gradRate_teacherData_graph
```

```{r}
StudentTeacherRatio_graph <-
  ggplot(gradRate_teacherData, aes(x = student_teacher_ratio, 
                                 y = percent_graduated )) + geom_point() + 
  theme_minimal() + 
  ggtitle("Graduate rate percentage vs. 
          Student teacher ratio") + geom_smooth(method=lm) 

StudentTeacherRatio_graph
```

# Student Discipline vs. Graduation Rate
```{r}


gradRate_discipline <- left_join(gradRate, discipline, by="District Code")
gradRate_discipline <- na.omit(gradRate_discipline)

gradRate_discipline <- gradRate_discipline %>%
  rename(
    'percent_graduated' = '% Graduated')

gradRate_discipline <- gradRate_discipline %>%
  rename(
    'percent_suspension' = '% In-School Suspension')

print(gradRate_discipline)
```

```{r}
gradRate_discipline_graph <-
  ggplot(gradRate_discipline, aes(x = percent_suspension, 
                                 y = percent_graduated )) + geom_point() + 
  theme_minimal() + 
  ggtitle("Graduate rate percentage vs. 
          Student discipline") + geom_smooth(method=lm) 

gradRate_discipline_graph
```

# Demographic vs. Graduation Rate
```{r}


gradRate_demographic <- left_join(gradRate, eduGen, by="District Code")
gradRate_demographic <- na.omit(gradRate_demographic)

gradRate_demographic <- gradRate_demographic %>%
  rename(
    'percent_graduated' = '% Graduated')

gradRate_demographic <- gradRate_demographic %>%
  rename(
    'African_American' = 'African American (#)')
gradRate_demographic <- gradRate_demographic %>%
  rename(
    'White' = 'White (#)')

gradRate_demographic <- gradRate_demographic %>%
  rename(
    'Asian' = 'Asian (#)')

gradRate_demographic <- gradRate_demographic %>%
  rename(
    'Hispanic' = 'Hispanic (#)')

gradRate_demographic <- gradRate_demographic %>%
  rename(
    'Native_American' = 'Native American (#)')

gradRate_demographic <- gradRate_demographic %>%
  rename(
    'Hawaiian_Pacific_Islander' = 'Native Hawaiian, Pacific Islander (#)')

gradRate_demographic <- gradRate_demographic %>%
  rename(
    'Multi_race_non_Hispanic' = 'Multi-Race,Non-Hispanic (#)')

gradRate_demographic_long <- pivot_longer(gradRate_demographic, cols=11:17, names_to = "Race", values_to = "Race_number")

print(gradRate_demographic_long)
```
```{r}
gradRate_demographic_graph <-
  ggplot(gradRate_demographic_long, aes(x = Race_number, 
                                 y = percent_graduated, color = Race)) + geom_point() + 
  theme_minimal() + 
  ggtitle("Graduate rate percentage vs. 
          Staff Demographic") + xlim(0, 1000)

gradRate_demographic_graph 
```

# Staffing Retention vs. Graduation Rate

```{r}

gradRate_staffingRetention <- left_join(gradRate, staffingRetention, by="District Code")
gradRate_staffingRetention <- na.omit(gradRate_staffingRetention)

print(gradRate_staffingRetention)
```

# Graduation Rate vs. Day missed

```{r}

print(daysMissed)



str_pad(daysMissed$District_code, 8, pad = "0")

gradRateRename <- gradRate %>%
  rename(
    'District_code' = 'District Code')

print(daysMissed)

```
gradRate_daysMissed <- left_join(gradRateRename, daysMissed, by="District_code")
gradRate_daysMissed <- na.omit(gradRate_daysMissed)

print(gradRate_daysMissed)

# SAT vs. Graduation Rate

```{r}

gradRate_SAT <- left_join(gradRate, sat, by="District Code")


gradRate_SAT <- gradRate_SAT %>%
  rename(
    'percent_graduated' = '% Graduated')

gradRate_SAT <- gradRate_SAT %>%
  rename(
    'tests_taken' = 'Tests Taken')

gradRate_SAT_long <- pivot_longer(gradRate_SAT, cols=12:14, names_to = "SAT_test_types", values_to = "SAT_test_scores")

gradRate_SAT_long <- gradRate_SAT_long[!(is.na(gradRate_SAT_long$SAT_test_scores)), ]

print(gradRate_SAT_long)
```

```{r}
gradRate_SAT_graph <-
  ggplot(gradRate_SAT_long, aes(x = SAT_test_scores, 
                                 y = percent_graduated, color = SAT_test_types)) + geom_point() + 
  theme_minimal() + 
  ggtitle("Graduate rate percentage vs. 
          SAT") 
gradRate_SAT_graph 
```

# Graduation plan Vs. Graduation rate

```{r}

gradRate_gradplan <- left_join(gradRate, plansforHSgrad, by="District Code")

gradRate_gradplan <- gradRate_gradplan %>%
  rename(
    'percent_graduated' = '% Graduated')

gradRate_gradplan_long <- pivot_longer(gradRate_gradplan, cols=11:20, names_to = "Plan_type", values_to = "Plan_percentage")

print(gradRate_gradplan_long)
```
```{r}
gradRate_gradPlan_graph <-
  ggplot(gradRate_gradplan_long, aes(x = Plan_percentage, 
                                 y = percent_graduated, color = Plan_type)) + geom_point() + 
  theme_minimal() + 
  ggtitle("Graduate rate percentage vs. 
          Plan after high school") 
gradRate_gradPlan_graph 
```














gradRate_SAT <- gradRate_SAT %>%
  rename(
    'tests_taken' = 'Tests Taken')

gradRate_SAT_long <- pivot_longer(gradRate_SAT, cols=12:14, names_to = "SAT_test_types", values_to = "SAT_test_scores")

gradRate_SAT_long <- gradRate_SAT_long[!(is.na(gradRate_SAT_long$SAT_test_scores)), ]






































